%% Harford User Manual, Viviano, Joseph David Apr. 2014
\documentclass[final,titlepage,letterpaper,oneside,12pt]{article}
\usepackage[usenames,dvipsnames]{xcolor} % colors
\usepackage{tikz}                       % graphs
\usepackage{caption}                    % figure captions
\usepackage[utf8]{inputenc}             % utf8 encoding compatibility
\usepackage{fourier}                    % fonts
\usepackage[T1]{fontenc}                % fonts
\usepackage{natbib}                     % bibliography
\usepackage[colorlinks=true, pdfstartview=FitV, linkcolor=MidnightBlue, citecolor=MidnightBlue,urlcolor=MidnightBlue]{hyperref}                % custom colors
\usetikzlibrary{arrows}                 % arrows
\usepackage{amsmath}                    % equations
\usepackage{epigraph}                   % quotation
\renewcommand\epigraphflush{flushright} % 
\renewcommand\epigraphsize{\normalsize} % 
\setlength\epigraphwidth{0.7\textwidth} % 
\definecolor{titlepagecolor}{cmyk}{1,.60,0,.40} % custom color

%% Title Page
\makeatletter                       
\def\printauthor{%                  
    {\normalsize \@author}}              
\makeatother
\author{%
    Joseph D. Viviano \\
    Department of Biology \\
    York University \\
    Toronto, ON, CA \\
    joseph.d.viviano@gmail.com
}

\newcommand\titlepagedecoration{%
\begin{tikzpicture}[remember picture,overlay,shorten >= -10pt]

\coordinate (aux1) at ([yshift=-15pt]current page.north east);
\coordinate (aux2) at ([yshift=-410pt]current page.north east);
\coordinate (aux3) at ([xshift=-4.5cm]current page.north east);
\coordinate (aux4) at ([yshift=-150pt]current page.north east);

\end{tikzpicture}%
}

\begin{document}
\begin{titlepage}

\noindent
\begin{center}
\begin{Huge}
Harford\par
\begin{small}
the\par
\end{small}
York University MRI Analysis Server\par
\end{Huge}
\end{center}
\vspace*{2cm}
\epigraph{Any sufficiently advanced technology is indistinguishable from magic.}%
{\textit{Hazards of Prophecy: The Failure of Imagination}\\ \textsc{Arthur C. Clarke}}
\null\vfill
\vspace*{1cm}
\noindent
\hfill
\begin{minipage}{0.55\linewidth}
    \begin{flushright}
        \printauthor
    \end{flushright}
\end{minipage}
%
\begin{minipage}{0.02\linewidth}
    \rule{1pt}{72pt}
\end{minipage}
\titlepagedecoration
\end{titlepage}

%% Document
\section{Introduction}

This server was built by a complete stranger to server-building, and I undoubtedly made some silly mistakes along the way. This document is my effort to crystallize these decisions for those who paid for it (henceforth \textit{the money}) and whoever takes over responsibility from me (henceforth \textit{the unlucky}). Hopefully this isn't my last job ever, and I still walk this earth. If that is true, feel free to contact me at any time \texttt{joseph.d.viviano@gmail.com} 

\section{Connecting}

To facilitate easy access to the server from your personal computer, on or off campus, we have configured three standard access methods.

\subsection{Samba}
\subsection{SSH}
\subsection{x2go}

Normally, you will want to view a full graphic-user interface while you interact with the server. This will allow you to easily view the data remotely, and make use of various programs not available with terminal-only access. To do this, we use x2go (\href{http://wiki.x2go.org/doku.php}{download x2go client here}). On Windows and Mac, you should be able to open x2go by clicking on the installed icon. On Linux, you may need to type \texttt{x2goclient} into the terminal.

\subsubsection{Troubleshooting x2go}

In some cases, you will receive an error involving the RSA key, and will not be allowed to log into the server. Your RSA keys are stored in \texttt{~/.ssh/known\_hosts}. You can inspect the contents of this file by typing \texttt{cat ~/.ssh/known\_hosts} into the terminal. If you find an entry that starts with \texttt{130.63.40.171}, remove it using either: \texttt{gedit ~/.ssh/known\_hosts} or \texttt{vi ~/.ssh/known\_hosts} (\href{http://glaciated.org/vi/}{here are some instructions for vi}). If you are having trouble doing this, you can simply reset the \texttt{known\_hosts} file by removing it with \texttt{rm ~/.ssh/known\_hosts} (this might affect other systems relying on RSA keys, but I will assume it won't if you are following these instructions). You should now be able to log into the server using x2go.

\subsection{Passwords}

Your default password can be changed to something more easily remembered by doing the following:
\begin{enumerate} \itemsep-2pt
    \item{SSH into the server using port \href{http://en.wikipedia.org/wiki/Looking_Glass_Studios}{45100}.}
    \begin{itemize} \itemsep-2pt
        \item{Harford uses a non-standard SSH port to prevent random attacks. This requires you to specify it at the command line with the following command: \texttt{ssh -p 45100 [your username]\@130.63.40.171}.}
    \end{itemize}
    \item{Use the \texttt{passwd} command.}
    \begin{itemize} \itemsep-2pt
        \item{If you type passwd at the command line, you will be prompted for your old annoying password, and asked to enter a new one.Your new password must satisfy the following conditions:}
        \begin{itemize} \itemsep-2pt
            \item{1 Uppercase letter}
            \item{1 lowercase letter}
            \item{1 number}
            \item{1 symbol (i.e., *, \_\#\$\%)}
            \item{12 characters length minimum.}
        \end{itemize}
        \item{This does not need to be hard to remember! Computer don't think like us.}
        \begin{itemize} \itemsep-2pt
            \item{\texttt{\#sQ\_} is easy for a computer to guess.}
            \item{\texttt{Br4in-fun-4eva!} is much harder.}
        \end{itemize}
    \end{itemize}
\end{enumerate}

Be sure to write your password down. If you forget your password, an administrator can reset it for you. If you \textit{are} an administrator, you should know you can change \texttt{username}'s password by dropping to \texttt{root} and typing \texttt{passwd [username]}.

Also, recall that samba and unix passwords are \textit{not} automatically synchronized. [This really should be fixed.]

\section{Hardware}

Harford is a \texttt{Thinkserver TS440 ($8 \times 3.5"$) HDD Hot-Swappable} server loaded with 4 HHD at the moment. Hard drive caddy part \texttt{\# 03X3969, FRU HS 3.5” HDD Tray V3.0}. It is loaded with 4 $\times$ Western Digital RED 4.0 TB 5400 HDD, configured in RAID 1+0. This currently leaves us with 8 TB of usable hard drive space, striped arrays, and on-site redundancy. It can easily be expanded to 2$\times$ that with an additional 4 drives.

\section{Server Architecture}
\subsection{Partitions}

At the moment, Harford consists of three partitions: 

\begin{enumerate}
    \item{\verb=/boot=: a 200 MB partition that is integral to the life of the server. This contains the \textit{Linux kernels} of the system -- the server's brainstem. Do not mess with this partition.}
    
    \item{\verb=/=: a 75 GB partition containing all user \texttt{/home} folders, software, and the operating system. This \textit{should} be large enough for indefinite expansion, if people don't store data in their home folders.}
    
    \item{\verb=/srv=: a 6.59 TB partition containing all of our data.}
\end{enumerate}


\begin{center}
\textbf{NB: this means that the lion's share of the disk space is found under \texttt{/srv} and therefore large files should \textit{always} be kept there!}
\end{center}

\subsection{Operating System}

I chose Ubuntu Server 12.04 LTS for its excellent support, modern features, and compatibility with the NeuroDebian project (\url{neuro.debian.net}). I named the server Harford after the Kubrickian hero. When I first installed the server, I was left only with a basic terminal. To get things normal-looking, I had to \texttt{sudo apt-get install} the following:

\noindent
\begin{small}
\begin{itemize} \itemsep-2pt
    \item{openssh-server}
    \item{xfce4}
    \item{kde-plasma-desktop}
    \item{synaptic}
    \item{lightdm-gtk-greeter}
    \item{jockey-gtk}
    \item{dmz-cursor-theme} 
    \item{xubuntu-icon-theme}
    \item{elementary-icon-theme}
\end{itemize}
\end{small}


This will produce a basic desktop environment for you to work in. I include both the \texttt{xfce4} and \texttt{kde} desktop environments, which each have their own merits and utility. They essentially control the graphic user interface of the server but do not differ substantially in actual usability.

A note on what these basic programs do. \texttt{openssh-server} enables SSH access through the terminal and is how people can control the server remotely. \texttt{xfce4} and \texttt{kde-plasma-desktop} are two alternative graphic user interfaces for the computer. \texttt{synaptic} allows one to probe the Internet for possible software installations and largely automates that process. Most of the software on the server can be installed and/or uninstalled using this simple program. \texttt{lightdm-gtk-greeter} presents the user with a login screen. The remaining installs are all graphic user interface niceties.

\subsection{Directories}
\subsection{Permissions}

The MRI data is separated into 3 main branches: RAW, WORKING, and ANALYSIS. These different tiers are more or less editable by various users of the server to strike a balance between data-security and usability. As a general rule, we also don't let unauthorized people look at data they aren't supposed to for both privacy and competitive reasons. These permissions are maintained by a \texttt{root}-owned nightly \texttt{CRON} job {\color{red}\texttt{maintain\_permissions.sh}}.

The \texttt{grandvizier} user should be used by the system administrators to preform various tasks without the need of dropping to \texttt{root}, which I am trying to discourage as it can be dangerous to spend too much time with so much power. If you can see a file owned by the \texttt{grandvizier}, that typically means it is being protected.  

\subsubsection{RAW}

These are DICOM files. Mostly used for archival purposes. Shouldn't be edited.

\begin{flushleft}
\texttt{owner = grandvizier, rwx \\
        group = staff      , r-x \\
        else  = ---} \\
\end{flushleft}

\subsubsection{WORKING}

These are the files manipulated by the pipeline code. Generally, these files should only be accessed by the pipeline and not manually. Right now, experiment specific group-wise permissions allow for you to go in and delete everything \textit{except} the input RAW data at the bottom of the WORKING tree. This allows you to `reset' problem subjects or whole users using the {\color{red}\texttt{cleanup\_X.sh}} programs, or manually remove problem files.

\begin{flushleft}

\texttt{owner = grandvizier, rwx \\
        group = [experiment], rwx (except inputs which are r-x) \\
        else  =            , ---} \\
\end{flushleft}

\subsubsection{ANALYSIS}

These folders are where people \textit{are} allowed to mess around. Generally, outputs from the pipeline can be copied into the ANALYSIS tree for manipulation. This is zero-risk as there is always an identical copy of the pre-processed files in the WORKING directory.

\begin{flushleft}
\texttt{owner = grandvizier, rwx \\
        group = [exrperiment], rwx \\
        else  =            , ---} \\
\end{flushleft}

\subsubsection{Privacy Notes}

The \texttt{grandvizier} user is special, and shared among \textit{the money} and \textit{the unlucky}. The \texttt{grandvizier} is also capable of  destroying millions of tax-payer dollars in a one-line command, so it shouldn't be used by anyone unless required.

All pipeline code will be run by the individuals within an experiment group -- and the code will work so long as the individual is permitted to interact with a given data set. This is safe because the pipeline code isn't editable by the users in the first place, so we can't do undue harm to our data by mistake.

\subsection{Security}
\subsection{Backup}
\section{Software \& Configuration}

The following is a list of the software installed on the server, and if you are lucky, it is even up-to-date.

\noindent
\begin{small}
\begin{itemize} \itemsep-2pt
    
    \item{\href{http://neuro.debian.net/}{NeuroDebian}:}
    \begin{itemize} \itemsep-2pt
        \item{\href{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/}{FSL 5.0.6}}
        \item{\href{http://www.mccauslandcenter.sc.edu/mricro/mricron/dcm2nii.html}{DICOM2NIFTI}}
        \item{\href{http://afni.nimh.nih.gov/afni/}{AFNI}}
        \item{\href{https://surfer.nmr.mgh.harvard.edu/}{FreeSurfer}}
        \item{\href{http://brainvis.wustl.edu/wiki/index.php/Caret:About}{Caret}}
        \item{\href{http://nipy.org/nibabel/}{Python: NiBabel}}
    \end{itemize}
    
    \item{General Computing:}
    \begin{itemize} \itemsep-2pt
        \item{\href{http://cran.r-project.org/}{R}}
        \item{\href{https://wiki.gnome.org/Apps/Gedit}{gedit}}
        \item{\href{https://launchpad.net/terminator}{Terminator}}
        \item{\href{https://www.libreoffice.org/}{LibreOffice}}
        \item{\href{https://wiki.gnome.org/Apps/Evince}{Evince}}
        \item{\href{https://wiki.gnome.org/Apps/Evince}{Firefox}}
        \item{\href{http://www.inkscape.org/en/}{Inkscape}}
        \item{\href{http://www.gimp.org/}{Gimp}}
        \item{\href{https://wiki.gnome.org/Apps/Dia}{Dia}}
        \item{\href{http://www.videolan.org/vlc/index.html}{VLC}}
        \item{\href{https://www.samba.org/}{Samba}}
    \end{itemize}
    
    \item{Python Packages:}
    \begin{itemize} \itemsep-2pt
        \item{\href{http://www.numpy.org/}{NumPy}}
        \item{\href{http://www.scipy.org/}{SciPy}} 
        \item{\href{http://pandas.pydata.org/}{Pandas}}
        \item{\href{http://matplotlib.org/}{matplotlib}}
        \item{\href{http://scikit-learn.org/stable/}{scikit-learn}}
        \item{\href{http://scikit-image.org/}{scikit-image}}
        \item{\href{http://networkx.github.io/}{NetworkX}}
        \item{\href{http://www.pymvpa.org/}{PyMVPA}}
        \item{\href{http://ipython.org/}{iPython}}
    \end{itemize}

    \item{Misc -- Installed Manually in \texttt{/opt}:}
    \begin{itemize} \itemsep-2pt
        \item{\href{https://gephi.org/}{Gephi}}
        \item{\href{http://www.mathworks.com/products/compiler/mcr/}{MATLAB Compiler Runtime}}
        \item{\href{http://afni.nimh.nih.gov/sscc/dglen/McRetroTS}{McRetro}}
    \end{itemize}
\end{itemize}
\end{small}

\subsection{MRI Tools}
\subsection{Programming Languages}
\section{Further Reading}

%% End of document. Go to reference list.
\newpage 
\bibliographystyle{apalike}
\bibliography{harford}
\end{document}